\newpage
\pagenumbering{Roman}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%                                中文摘要
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\centerline{\large \zihao{3}\heiti \hbox{摘要}} \vskip 0.3cm
车载组合导航是当前智能车辆必不可少的一项重要功能。为满足自动辅助驾驶系统对小型化低成本低功耗导航系统的需求，本课题基于微电子机械系统(MEMS)惯性传感器和全球定位系统，完成了一款MEMS惯导/GNSS组合导航模块的设计。

本文着重理解分析了组合导航的基本原理，建立MEMS惯导姿态、速度和位置误差模型，推导了卡尔曼滤波的状态方程和观测方程。系统采用位置/速度松组合导航算法和卡尔曼滤波器完成导航系统数据的信息融合，研究过程主要可以可分成硬件设计和软件设计。主要研究内容如下：

1、导航模块以ARM微处理器、MEMS惯性器件、卫星接收机等为主要组成部分，详细给出系统硬件结构框架、硬件选型方案及接口设计。最后完成硬件原理图和电路板实物制作。

2、以MEMS惯导15维导航参数误差作为状态量，建立组合系统数学模型。在硬件实验平台的基础上，提出了系统相应的软件设计方案，并编写组合导航算法软件，实现了组合系统的多传感器通信，数据的融合解算、导航信息的存储等功能。

3、设计了MEMS惯导/GNSS组合系统的实时监控软件，可以实现导航模式选择、运行轨迹显示、历史数据加载、数据的保存等功能，可以直观地显示车辆的运行状态。

最后，对所设计的组合导航模块进行车载实验，结果表明本设计的MEMS惯导/GNSS组合系统完成了实时导航系统的基本功能，且位置误差约$5\sim7m$，速率误差约$0.2\sim0.5m/s$，姿态误差约为$8^\circ$，满足车载组合系统小型化、低成本和实时性的设计要求。

\vskip 0.3cm \textbf{关键词}: 小型化；MEMS惯导；车载组合导航；卡尔曼滤波

\newpage

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%                                英文摘要
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

$\\$ \centerline{\large \zihao{3}\heiti\hbox{Abstract}}

In-vehicle Navigation System is fundamental to smart automobile.To meet Auto-pilot System's demand of a miniaturized navigation system with low cost and low energy consumption,A diminutive MEMS(Micro-Electro-Mechanical-Systems)-SINS/GNSS integrated navigation module was designed.

This dissertation builds on principles of in-vehicle integrated navigation, establishes error correction models of attitude, speed and location, and deduces equations of state and measurement of Kalman Filtering.  It then employs the location/speed integrated navigation algorithm and Kalman Filtering to process data. The design work involved in the system includes designs of hardware and software,with details as follows. 

(1)The navigation module consists of ARM microprocessor, MEMS inertial devices,a satellite navigation module,etc.The hardware framework, chip selection and communication interface design are clearly illustrated, and the hardware schematic diagram and prototype are completed.

(2)The algorithm establishes a mathematical model of the integrated system, using 15 navigation parameters of MEMS INS as status values. Based on the hardware experiment platform, corresponding software designs are proposed, and the algorithm software of the integrated navigation is programmed. Functions including multi-sensor communications, integrated settlement and navigation information storage are achieved.

(3)A real-time monitoring software is designed for the MEMS INS/GNSS system to bring about functions including mode selection, trajectory display and data storage. This helps a direct display of automobiles’ operating conditions.

An on-site experiment has been carried out, showing the designed MEMS INS/GNSS integrated system fully performs all functions of a real-time navigation system, with the position error of $5\sim7m$, the velocity error of $0.2\sim0.5m/s$ and the attitude error of $8^\circ$. The result stands firmly to meet the demand of a miniaturized real-time in-vehicle navigation system with low cost.

\textbf{Keywords}:miniaturized;MEMS INS;Vehicle integrated navigation;Kalman filtering

\newpage
